Data_Analysis_exp7

### 2018-07-04 Results/preproc/MS_ddm_preproc.rmd
revising exp1's code

This script is aimed at making the analysis of the experiment 7 reproducible. We analyzezd the data using the traditional ways (d prime/Accuracy and mean of RT)

```{r Initializing, include=FALSE}
source('Initial.r')
curDir <- getwd()   # directory for preprocessing
rootDir <- gsub('.{7}$', '', curDir)
#rawDir <- paste(curDir,'/data/',sep = '')
traDir <- paste(rootDir,'Traditional_Analysis',sep = '')
ddmDir <- paste(rootDir,'HDDM',sep = '')
```

```{r loadingData_e7,echo=FALSE,results='hide'}
df.L <- read.csv("df.Learn.csv",header = TRUE, sep = ',', stringsAsFactors=FALSE) # data for matching task
df.T <- read.csv("df.Test.csv",header = TRUE, sep = ',', stringsAsFactors=FALSE)  # data for categorization task

# render the data in numeric format for future analysis
cols.num <- c("Age","Block","Bin","Trial","RT","Accuracy")
df.L[cols.num] <- sapply(df.L[cols.num], as.numeric)  
df.T[cols.num] <- sapply(df.T[cols.num], as.numeric)

# rename columns
colnames(df.L)[colnames(df.L)=="Accuracy"]  <- "ACC"
colnames(df.T)[colnames(df.T)=="Accuracy"]  <- "ACC"
colnames(df.L)[colnames(df.L)=="SubjectID"] <- "Subject"
colnames(df.T)[colnames(df.T)=="SubjectID"] <- "Subject"

df.L$Subject <- factor(df.L$Subject)
df.T$Subject <- factor(df.T$Subject)

# Change the name for different levels of independent variables
df.L$Morality[df.L$Morality == 'moral']   <- 'Good'
df.L$Morality[df.L$Morality == 'immoral'] <- 'Bad'
df.L$Identity[df.L$Identity == 'self']    <- 'Self'
df.L$Identity[df.L$Identity == 'other']   <- 'Other'

# make the variables in a specified order
df.L$Morality <- factor(df.L$Morality, levels = c("Good","Bad"))    
df.L$Identity <- factor(df.L$Identity, levels = c("Self","Other"))  
df.L$Match    <- factor(df.L$Match,    levels = c("match","nonmatch"))

# Change the name for different levels of independent variables
df.T$Morality[df.T$Morality == 'moral']   <- 'Good'
df.T$Morality[df.T$Morality == 'immoral'] <- 'Bad'
df.T$Identity[df.T$Identity == 'self']    <- 'Self'
df.T$Identity[df.T$Identity == 'other']   <- 'Other'
df.T$Task[df.T$Task == 'Self']            <- 'ID'
df.T$Task[df.T$Task == 'self']            <- 'Id_task'
df.T$Task[df.T$Task == 'moral']           <- 'Valence_task'

# make the variables in a specified order
df.T$Morality <- factor(df.T$Morality, levels=c("Good","Bad"))    
df.T$Identity <- factor(df.T$Identity, levels=c("Self","Other"))
```

```{r clean the data_e7,echo=FALSE,results='hide',warning=FALSE, message=FALSE}
df.L1 <- df.L
df.T1 <- df.T
df.L1$RT <- df.L1$RT * 1000 # transfer from seconds to min seconds
df.T1$RT <- df.T1$RT * 1000 # careful about the scale of time when using HDDM or ex-Gaussian

# no response equal to wrong
df.L1$ACC[df.L1$ACC == -1] <- 0

# excluded data from subjects whose trials number are not normal
procFailureSub <- c("7","8","2027","7035")
df.L1 <- df.L1[!(df.L1$Subject %in% procFailureSub),]
df.T1 <- df.T1[!(df.T1$Subject %in% procFailureSub),]
df.T1 <- df.T1[df.T1$Task != 'importance',]  # exclude the results from importance task

# delete the practicing trials (first 48 trials)
subNo <- unique(df.L1$Subject)
for (subj in subNo) {
        if (exists('df.L1.fm')){
                df.tmp <- df.L1[df.L1$Subject == subj,]
                df.tmp <- df.tmp[49:nrow(df.tmp),]
                df.L1.fm <- rbind(df.L1.fm,df.tmp)
        } else {
                df.L1.fm <- df.L1[df.L1$Subject == subj,]
                df.L1.fm <- df.L1.fm[49:nrow(df.L1.fm),]
        }
}

rm(subNo)

# calculate the overall accuracy for matching task
df.L1.acc.g <-  ddply(df.L1.fm,.(Subject), summarise,
                    N = length(ACC),
                    countN = sum(ACC),
                    ACC = sum(ACC)/length(ACC))

# calculate the overall accuracy for categorziation task
df.T1.acc.g <-  ddply(df.T1,.(Subject), summarise,
                    N = length(ACC),
                    countN = sum(ACC),
                    ACC = sum(ACC)/length(ACC))

excld.sub   <- df.L1.acc.g$Subject[df.L1.acc.g$ACC < 0.5]      # less 50% accuracy in matching task
excld.sub.T <- df.T1.acc.g$Subject[df.T1.acc.g$ACC < 0.5]      # less than 50% accuracy in categorization task
df.L1.valid <- df.L1.fm[!(df.L1.fm$Subject %in% excld.sub),]   # exclude the invalid subjects
df.T1.valid <- df.T1[!(df.T1$Subject %in% excld.sub),]
df.T1.valid <- df.T1.valid[!(df.T1.valid$Subject %in% excld.sub.T),]

# check the number of participants are correct
length(unique(df.L1.valid$Subject)) + length(excld.sub) == length(unique(df.L1$Subject))
# excld.trials3 <- excld.trials[!(excld.trials$Subject %in% excld.sub),]
excld.trials.L <- df.L1.valid[df.L1.valid$RT <= 200,]
df.L1.V        <- df.L1.valid[!(df.L1.valid$RT <= 200),]  # valid trial data for match task
excld.trials.T <- df.T1.valid[df.T1.valid$RT <= 200,]
df.T1.V        <- df.T1.valid[!(df.T1.valid$RT <= 200),]  # valid trial data for categorization task
nrow(excld.trials.T) + nrow(df.T1.V) == nrow(df.T1.valid) # if true, everything is ok

## Basic information of the data ####
df.L1.T.basic <- df.L1[!duplicated(df.L1$Subject), 1:4]
numT.subj <- nrow(df.L1.T.basic)
numT.female <- sum(df.L1.T.basic$Gender == 'female');
numT.male <- sum(df.L1.T.basic$Gender == 'male');
ageT.mean <- round(mean(df.L1.T.basic$Age),2);
ageT.std <- round(sd(df.L1.T.basic$Age),2);
num.excld.sub <- length(unique(excld.sub))
num.excld.sub.T <- length(unique(excld.sub.T))

# valide data for matching task
df.L1.V.basic <- df.L1.V[!duplicated(df.L1.V$Subject), 1:4]
numV.female <- sum(df.L1.V.basic$Gender == 'female');
numV.male <- sum(df.L1.V.basic$Gender == 'male');
ageV.mean <- round(mean(df.L1.V.basic$Age),2);
ageV.std <- round(sd(df.L1.V.basic$Age),2);
ratio.excld.trials.L <- nrow(excld.trials.L)/nrow(df.L1.valid)
ratio.excld.trials.T <- nrow(excld.trials.T)/nrow(df.T1.valid)

# valid data for categorization task
df.T1.V.basic <- df.T1.V[!duplicated(df.T1.V$Subject), 1:4]
numV.female.T <- sum(df.T1.V.basic$Gender == 'female');
numV.male.T <- sum(df.T1.V.basic$Gender == 'male');
ageV.mean.T <- round(mean(df.T1.V.basic$Age),2);
ageV.std.T <- round(sd(df.T1.V.basic$Age),2);
ratio.excld.trials.T <- nrow(excld.trials.T)/nrow(df.T1.valid)

# get the data file for hddm analysis

df.L.ddm.v <- df.L[!(df.L$Subject %in% c(procFailureSub,excld.sub)),]   # exclude the invalid subjects

df.T.ddm.v <- df.T[!(df.T$Subject %in% c(procFailureSub,excld.sub)),]
df.T.ddm.v <- df.T.ddm.v[!(df.T.ddm.v$Subject %in% excld.sub.T),]
df.T.ddm.v <- df.T[df.T$Task != 'importance',]  # exclude the results from importance task

#df.L1 <- df.L1[!(df.L1$Subject %in% procFailureSub),]
#df.T1 <- df.T1[!(df.T1$Subject %in% procFailureSub),]

#df.L1.valid <- df.L1.fm[!(df.L1.fm$Subject %in% excld.sub),]
#df.T1.valid <- df.T1[!(df.T1$Subject %in% excld.sub),]
#df.T1.valid <- df.T1.valid[!(df.T1.valid$Subject %in% excld.sub.T),]

df.L.hddm <- subset(df.L.ddm.v, ACC ==1 | ACC == 0)  # remove non-response trials or pressing wrong key
df.T.hddm <- subset(df.T.ddm.v, ACC ==1 | ACC == 0)

df.L.hddm_m <- subset(df.L.hddm, Match == 'match') # select data
df.L.hddm_m <- df.L.hddm_m[,c('Subject','Morality','Identity','RT','ACC')] # select column
colnames(df.L.hddm_m) <- c("subj_idx","val",'id','rt','response')     # change column name

df.L.hddm_nm <- subset(df.L.hddm, Match == 'mismatch')
df.L.hddm_nm <- df.L.hddm_nm[,c('Subject','Morality','Identity','RT','ACC')]
colnames(df.L.hddm_nm) <- c("subj_idx","val",'id','rt','response')

df.T.hddm_val <- subset(df.T.hddm, Task == 'Val')
df.T.hddm_val <- df.T.hddm_val[,c('Subject','Morality','Identity','RT','ACC')]
colnames(df.T.hddm_val) <- c("subj_idx","val",'id','rt','response')

df.T.hddm_id <- subset(df.T.hddm, Task == 'Id')
df.T.hddm_id <- df.T.hddm_id[,c('Subject','Morality','Identity','RT','ACC')]
colnames(df.T.hddm_id) <- c("subj_idx","val",'id','rt','response')

setwd(ddmDir)
write.csv(df.L.hddm_m,'exp7_rep_match_hddm.csv',row.names = F)
write.csv(df.L.hddm_nm,'exp7_rep_mismatch_hddm.csv',row.names = F)
write.csv(df.T.hddm_val,'exp7_rep_categ_val_hddm.csv',row.names = F)
write.csv(df.T.hddm_id,'exp7_rep_categ_id_hddm.csv',row.names = F)
rm('df.L.hddm','df.T.hddm','df.L.hddm_m','df.L.hddm_nm','df.T.hddm_val','df.T.hddm_id')
setwd(curDir)
```
## Participants
`r numT.subj+length(procFailureSub)` college students (`r numT.female` female, age: `r ageT.mean` $\pm$ `r ageT.std`) participated in experiment 7. All partcipants were right handed, and all had normal or corrected-to-normal vision. Informed consent was obtained from all partcipants prior to the experiment according to procedure approved by a local ethics committee. Data of `r length(procFailureSub)` participants were excluded from matching task because of the procedura failure, and `r num.excld.sub` of the participants data were excluded from the analysis for matching task and categorization task because of less than 50% overall accuracy for the matching task, leaving `r numT.subj - num.excld.sub` participants (`r numV.female` female, age: `r ageV.mean` $\pm$ `r ageV.std` years) for matching task. `r num.excld.sub.T` addtional participants's data in categorization task were excluded becasue of less than 50% accuracy for the categorization task,leaving `r numT.subj - num.excld.sub - num.excld.sub.T` participants (`r numV.female.T` female, age: `r ageV.mean.T` $\pm$ `r ageV.std.T` years) for categorization task.

@@

### Results of Learning phase
#### Analaysis of d prime
```{r analyzing for d prime_e7_match, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
## calculate the d prime for each condition
# calculate the number of hit,CR,miss or FA 
df.L1.V$sdt <- NA
for (i in 1:nrow(df.L1.V)){
        if (df.L1.V$ACC[i] == 1 & df.L1.V$Match[i] == "match" & !is.na(df.L1.V$ResponseKey[i])){
                df.L1.V$sdt[i] <- "hit"
        } else if (df.L1.V$ACC[i] == 1 & df.L1.V$Match[i] == "nonmatch" & !is.na(df.L1.V$ResponseKey[i])){
                df.L1.V$sdt[i] <- "CR"
        } else if (df.L1.V$ACC[i] == 0 & df.L1.V$Match[i] == "match" & !is.na(df.L1.V$ResponseKey[i])){
                df.L1.V$sdt[i] <- "miss"
        } else if (df.L1.V$ACC[i] == 0 & df.L1.V$Match[i] == "nonmatch" & !is.na(df.L1.V$ResponseKey[i])){
                df.L1.V$sdt[i] <- "FA"
        }
}

# calculate the number of each for each condition
df.L1.V.SDT <-  ddply(df.L1.V,.(Subject,Morality, Identity,sdt), summarise, N = length(sdt))
df.L1.V.SDT <- df.L1.V.SDT[complete.cases(df.L1.V.SDT$sdt),] # exclude NA, which represents no-response trials

# long format to wide
df.L1.V.SDT_w <- dcast(df.L1.V.SDT, Subject + Morality + Identity ~ sdt,value.var = "N")
df.L1.V.SDT_w$miss[is.na(df.L1.V.SDT_w$miss)] <- 0 # transfer NA to 0
df.L1.V.SDT_w$FA[is.na(df.L1.V.SDT_w$FA)]     <- 0
df.L1.V.SDT_w$hitR <- df.L1.V.SDT_w$hit/(df.L1.V.SDT_w$hit + df.L1.V.SDT_w$miss)
df.L1.V.SDT_w$faR <- df.L1.V.SDT_w$FA/(df.L1.V.SDT_w$FA + df.L1.V.SDT_w$CR)

# standardized way to deal with the extreme values
for (i in 1:nrow(df.L1.V.SDT_w)){
        if (df.L1.V.SDT_w$hitR[i] == 1){
                df.L1.V.SDT_w$hitR[i] <- 1 - 1/(2*(df.L1.V.SDT_w$hit[i] + df.L1.V.SDT_w$miss[i]))
        }
}

for (i in 1:nrow(df.L1.V.SDT_w)){
        if (df.L1.V.SDT_w$faR[i] == 0){
                df.L1.V.SDT_w$faR[i] <- 1/(2*(df.L1.V.SDT_w$FA[i] + df.L1.V.SDT_w$CR[i]))
        }
}

# calculate the d prime for each condition
df.L1.V.SDT_w$dprime <- mapply(dprime,df.L1.V.SDT_w$hitR,df.L1.V.SDT_w$faR)

# transfor from long format to wide format
df.L1.V.SDT_ww <- dcast(df.L1.V.SDT_w, Subject ~ Identity + Morality ,value.var = "dprime")

# anova for d prime with 2*2 design
e7_L.d_anova <- ezANOVA(df.L1.V.SDT_w,dv = dprime, wid = Subject, within=.(Morality,Identity), type=3)

df.L1.V.SDT_w$Subject <- factor(df.L1.V.SDT_w$Subject)  # make the subject "factor"

# using Bayesfactor analysis (which is not reported in the document)
e7_L.dprime_anova_bf  <- anovaBF(dprime ~ Morality*Identity + Subject, data = df.L1.V.SDT_w, whichRandom = "Subject")

# Good self vs Bad self
e7_L.d.t.good_bad_slf <- t.test(df.L1.V.SDT_ww$Self_Good,df.L1.V.SDT_ww$Self_Bad,paired = TRUE)     # t-test
df.L1.V.SDT_ww$good_bad_slf <- df.L1.V.SDT_ww$Self_Good - df.L1.V.SDT_ww$Self_Bad                   # calculate the differences
e7_L.d.t.good_bad_slf.CI <- bootES(df.L1.V.SDT_ww$good_bad_slf,R = 20000, effect.type = "cohens.d") # calculate the cI using BootES

# give new names for each key information, for later use in the markdown
e7_L.d.tvalue.good_bad_slf <- round(as.numeric(e7_L.d.t.good_bad_slf[[1]]),3)
e7_L.d.df.L1.good_bad_slf <- as.numeric(e7_L.d.t.good_bad_slf[[2]])
e7_L.d.pvalue.good_bad_slf.adj <- p.adjust(as.numeric(e7_L.d.t.good_bad_slf[[3]]),"bonferroni",4)
e7_L.d.cohens.good_bad_slf <- round(e7_L.d.t.good_bad_slf.CI[[1]],4) 
e7_L.d.CI.L.good_bad_slf <- round(e7_L.d.t.good_bad_slf.CI[[12]][1],4)
e7_L.d.CI.H.good_bad_slf <- round(e7_L.d.t.good_bad_slf.CI[[12]][2],4)

# Good other vs. Bad other
e7_L.d.t.good_bad_oth <- t.test(df.L1.V.SDT_ww$Other_Good,df.L1.V.SDT_ww$Other_Bad,paired = TRUE)
df.L1.V.SDT_ww$good_bad_oth <- df.L1.V.SDT_ww$Other_Good - df.L1.V.SDT_ww$Other_Bad
e7_L.d.t.good_bad_oth.CI <- bootES(df.L1.V.SDT_ww$good_bad_oth,R = 20000, effect.type = "cohens.d")

e7_L.d.tvalue.good_bad_oth  <- round(as.numeric(e7_L.d.t.good_bad_oth[[1]]),3)
e7_L.d.df.L1.good_bad_oth  <- as.numeric(e7_L.d.t.good_bad_oth[[2]])
e7_L.d.pvalue.good_bad_oth.adj <- p.adjust(as.numeric(e7_L.d.t.good_bad_oth[[3]]),"bonferroni",4)
e7_L.d.cohens.good_bad_oth <- round(e7_L.d.t.good_bad_oth.CI[[1]],4) 
e7_L.d.CI.L.good_bad_oth <- round(e7_L.d.t.good_bad_oth.CI[[12]][1],4)
e7_L.d.CI.H.good_bad_oth <- round(e7_L.d.t.good_bad_oth.CI[[12]][2],4)

# Good self vs Good other
e7_L.d.t.slf_oth_good <- t.test(df.L1.V.SDT_ww$Self_Good,df.L1.V.SDT_ww$Other_Good,paired = TRUE)
df.L1.V.SDT_ww$slf_oth_good <- df.L1.V.SDT_ww$Self_Good - df.L1.V.SDT_ww$Other_Good
e7_L.d.t.slf_oth_good.CI <- bootES(df.L1.V.SDT_ww$slf_oth_good,R = 20000, effect.type = "cohens.d")

>>>>

This script is aimed at making the analysis of the moral self experiment (DDM) reproducible. We analyzezd the data using the traditional ways (d prime/Accuracy and mean of RT)

```{r Initializing, include=FALSE}
source('Initial.r')
curDir  <- dirname(rstudioapi::getSourceEditorContext()$path)   # directory for preprocessing
rootDir <- gsub('.{7}$', '', curDir)
#rawDir <- paste(curDir,'/data/',sep = '')
traDir <- paste(rootDir,'Traditional_Analysis',sep = '')
ddmDir <- paste(rootDir,'HDDM',sep = '')
```

```{r loadingData_e7,echo=FALSE,results='hide'}
df.M <- read.csv("MS_matchTask_raw.csv",header = TRUE, sep = ',', stringsAsFactors=FALSE) # data for matching task
df.C <- read.csv("MS_categTask_raw.csv",header = TRUE, sep = ',', stringsAsFactors=FALSE)  # data for categorization task

# make the variables in a specified order
df.M$Morality <- factor(df.M$Morality, levels = c("Good","Bad"))    
df.M$Identity <- factor(df.M$Identity, levels = c("Self","Other"))  
df.M$Match    <- factor(df.M$Match,    levels = c("match","nonmatch"))

# make the variables in a specified order
df.C$Morality <- factor(df.C$Morality, levels=c("Good","Bad"))    
df.C$Identity <- factor(df.C$Identity, levels=c("Self","Other"))

# Clear data for all following analysis

# Exclude Subject, criterion 1: procedure failure
excldSub1 <- c("7","8","2027","7035")
df.M <- df.M[!(df.M$Subject %in% excldSub1),]
df.C <- df.C[!(df.C$Subject %in% excldSub1),]

# exclude trials, criterio 1: no importance-based trials
df.C <- df.C[df.C$Task != 'importance',]  # exclude the results from importance task, should be 17856 rows

# exclude trials, criterio 2: practicing trials in matching task (first 48 trials)
subNo <- unique(df.M$Subject)
for (subj in subNo) {
        if (exists('df.M.fm')){
                df.tmp <- df.M[df.M$Subject == subj,]
                df.tmp <- df.tmp[49:nrow(df.tmp),]
                df.M.fm <- rbind(df.M.fm,df.tmp)
        } else {
                df.M.fm <- df.M[df.M$Subject == subj,]
                df.M.fm <- df.M.fm[49:nrow(df.M.fm),]
        }
}   # df.M.fm should be 14880 rows
df.M <- df.M.fm  # all the experimental trials
rm(subNo,df.M.fm,df.tmp)

# exlude subject criterion 2: less than 50% overall accuracy
# change no response to error
df.M.tmp <- df.M
df.M.tmp$ACC[df.M.tmp$ACC == -1] <- 0
df.C.tmp <- df.C
df.C.tmp$ACC[df.C.tmp$ACC == -1] <- 0

# calculate the overall accuracy for matching task
df.M.acc.g <-  ddply(df.M.tmp,.(Subject), summarise,
                    N = length(ACC),
                    countN = sum(ACC),
                    ACC = sum(ACC)/length(ACC))

# calculate the overall accuracy for categorziation task
df.C.acc.g <-  ddply(df.C.tmp,.(Subject), summarise,
                    N = length(ACC),
                    countN = sum(ACC),
                    ACC = sum(ACC)/length(ACC))

excldSub2.M <- df.M.acc.g$Subject[df.M.acc.g$ACC < 0.5]      # less 50% accuracy in matching task
excldSub2.C <- df.C.acc.g$Subject[df.C.acc.g$ACC < 0.5]      # less than 50% accuracy in categorization task
df.M.valid <- df.M[!(df.M$Subject %in% excldSub2.M),]        # exclude the invalid subjects
df.C.valid <- df.C[!(df.C$Subject %in% excldSub2.C),]
df.C.valid <- df.C.valid[!(df.C.valid$Subject %in% excldSub2.C),]
rm(df.M,df.C,df.M.tmp,df.C.tmp)
```

```{r clean the data_e7,echo=FALSE,results='hide',warning=FALSE, message=FALSE}
# clean data for traditional analysis
df.M1 <- df.M.valid
df.C1 <- df.C.valid
df.M1$RT <- df.M1$RT * 1000 # transfer from seconds to min seconds
df.C1$RT <- df.C1$RT * 1000 # careful about the scale of time when using HDDM or ex-Gaussian

# no response equal to wrong
df.M1$ACC[df.M1$ACC == -1] <- 0
df.C1$ACC[df.C1$ACC == -1] <- 0

# excld.trials, criterion 2: correct response within 200 ms
excld.trials.M <- df.M1[df.M1$RT <= 200,]
df.M1.V        <- df.M1[!(df.M1$RT <= 200),]  # valid trial data for match task
excld.trials.C <- df.C1[df.C1$RT <= 200,]
df.C1.V        <- df.C1[!(df.C1$RT <= 200),]  # valid trial data for categorization task
nrow(excld.trials.C) + nrow(df.C1.V) == nrow(df.C1) # if true, everything is ok

## Basic information of the data ####
df.M1.T.basic <- df.M1[!duplicated(df.M1$Subject), 1:4]
numT.subj <- nrow(df.M1.T.basic)
numT.female <- sum(df.M1.T.basic$Gender == 'female');
numT.male <- sum(df.M1.T.basic$Gender == 'male');
ageT.mean <- round(mean(df.M1.T.basic$Age),2);
ageT.std <- round(sd(df.M1.T.basic$Age),2);
num.excldSub2.M <- length(unique(excldSub2.M))
num.excldSub2.C <- length(unique(excldSub2.C))

# valide data for matching task
df.M1.V.basic <- df.M1.V[!duplicated(df.M1.V$Subject), 1:4]
numV.female <- sum(df.M1.V.basic$Gender == 'female');
numV.male <- sum(df.M1.V.basic$Gender == 'male');
ageV.mean <- round(mean(df.M1.V.basic$Age),2);
ageV.std <- round(sd(df.M1.V.basic$Age),2);
ratio.excld.trials.M <- nrow(excld.trials.M)/nrow(df.M1)

# valid data for categorization task
df.C1.V.basic <- df.C1.V[!duplicated(df.C1.V$Subject), 1:4]
numV.female.C <- sum(df.C1.V.basic$Gender == 'female');
numV.male.C <- sum(df.C1.V.basic$Gender == 'male');
ageV.mean.C <- round(mean(df.C1.V.basic$Age),2);
ageV.std.C <- round(sd(df.C1.V.basic$Age),2);
ratio.excld.trials.C <- nrow(excld.trials.C)/nrow(df.C1)

rm(df.M1,df.C1) # remove the no longer used varibles.

# get the data file for hddm analysis
df.M.ddm.v <- df.M.valid   # exclude the invalid subjects
df.C.ddm.v <- df.C.valid

# exclusion trials, criterion: trials without response or wrong key
# Note: we didn't remove the correct response less than 200 ms
df.M.hddm <- subset(df.M.ddm.v, ACC ==1 | ACC == 0)  
df.C.hddm <- subset(df.C.ddm.v, ACC ==1 | ACC == 0)

df.M.hddm_m <- subset(df.M.hddm, Match == 'match') # select data
df.M.hddm_m <- df.M.hddm_m[,c('Subject','Morality','Identity','RT','ACC')] # select column
colnames(df.M.hddm_m) <- c("subj_idx","val",'id','rt','response')     # change column name

df.M.hddm_nm <- subset(df.M.hddm, Match == 'nonmatch')
df.M.hddm_nm <- df.M.hddm_nm[,c('Subject','Morality','Identity','RT','ACC')]
colnames(df.M.hddm_nm) <- c("subj_idx","val",'id','rt','response')

df.C.hddm_val <- subset(df.C.hddm, Task == 'Val')
df.C.hddm_val <- df.C.hddm_val[,c('Subject','Morality','Identity','RT','ACC')]
colnames(df.C.hddm_val) <- c("subj_idx","val",'id','rt','response')

df.C.hddm_id <- subset(df.C.hddm, Task == 'Id')
df.C.hddm_id <- df.C.hddm_id[,c('Subject','Morality','Identity','RT','ACC')]
colnames(df.C.hddm_id) <- c("subj_idx","val",'id','rt','response')

setwd(ddmDir)
write.csv(df.M.hddm_m,'MS_match_hddm.csv',row.names = F)
write.csv(df.M.hddm_nm,'MS_mismatch_hddm.csv',row.names = F)
write.csv(df.C.hddm_val,'MS_categ_val_hddm.csv',row.names = F)
write.csv(df.C.hddm_id,'MS_categ_id_hddm.csv',row.names = F)
setwd(curDir)
rm('df.M.hddm','df.C.hddm','df.M.hddm_m','df.M.hddm_nm','df.C.hddm_val','df.C.hddm_id')
```
## Participants
`r numT.subj+length(excldSub1)` college students (`r numT.female` female, age: `r ageT.mean` $\pm$ `r ageT.std`) participated in experiment 7. All partcipants were right handed, and all had normal or corrected-to-normal vision. Informed consent was obtained from all partcipants prior to the experiment according to procedure approved by a local ethics committee. Data of `r length(excldSub1)` participants were excluded from matching task because of the procedura failure, and `r num.excld.sub` of the participants data were excluded from the analysis for matching task and categorization task because of less than 50% overall accuracy for the matching task, leaving `r numT.subj - num.excld.sub` participants (`r numV.female` female, age: `r ageV.mean` $\pm$ `r ageV.std` years) for matching task. `r num.excld.sub.T` addtional participants's data in categorization task were excluded becasue of less than 50% accuracy for the categorization task,leaving `r numT.subj - num.excld.sub - num.excld.sub.T` participants (`r numV.female.T` female, age: `r ageV.mean.T` $\pm$ `r ageV.std.T` years) for categorization task.

@@

### Results of Learning phase
#### Analaysis of d prime
```{r analyzing for d prime_e7_match, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
## calculate the d prime for each condition
# calculate the number of hit,CR,miss or FA 
df.M1.V$sdt <- NA
for (i in 1:nrow(df.M1.V)){
        if (df.M1.V$ACC[i] == 1 & df.M1.V$Match[i] == "match" & !is.na(df.M1.V$ResponseKey[i])){
                df.M1.V$sdt[i] <- "hit"
        } else if (df.M1.V$ACC[i] == 1 & df.M1.V$Match[i] == "nonmatch" & !is.na(df.M1.V$ResponseKey[i])){
                df.M1.V$sdt[i] <- "CR"
        } else if (df.M1.V$ACC[i] == 0 & df.M1.V$Match[i] == "match" & !is.na(df.M1.V$ResponseKey[i])){
                df.M1.V$sdt[i] <- "miss"
        } else if (df.M1.V$ACC[i] == 0 & df.M1.V$Match[i] == "nonmatch" & !is.na(df.M1.V$ResponseKey[i])){
                df.M1.V$sdt[i] <- "FA"
        }
}

# calculate the number of each for each condition
df.M1.V.SDT <-  ddply(df.M1.V,.(Subject,Morality, Identity,sdt), summarise, N = length(sdt))
df.M1.V.SDT <- df.M1.V.SDT[complete.cases(df.M1.V.SDT$sdt),] # exclude NA, which represents no-response trials

# long format to wide
df.M1.V.SDT_w <- dcast(df.M1.V.SDT, Subject + Morality + Identity ~ sdt,value.var = "N")
df.M1.V.SDT_w$miss[is.na(df.M1.V.SDT_w$miss)] <- 0 # transfer NA to 0
df.M1.V.SDT_w$FA[is.na(df.M1.V.SDT_w$FA)]     <- 0
df.M1.V.SDT_w$hitR <- df.M1.V.SDT_w$hit/(df.M1.V.SDT_w$hit + df.M1.V.SDT_w$miss)
df.M1.V.SDT_w$faR <- df.M1.V.SDT_w$FA/(df.M1.V.SDT_w$FA + df.M1.V.SDT_w$CR)

# standardized way to deal with the extreme values
for (i in 1:nrow(df.M1.V.SDT_w)){
        if (df.M1.V.SDT_w$hitR[i] == 1){
                df.M1.V.SDT_w$hitR[i] <- 1 - 1/(2*(df.M1.V.SDT_w$hit[i] + df.M1.V.SDT_w$miss[i]))
        }
}

for (i in 1:nrow(df.M1.V.SDT_w)){
        if (df.M1.V.SDT_w$faR[i] == 0){
                df.M1.V.SDT_w$faR[i] <- 1/(2*(df.M1.V.SDT_w$FA[i] + df.M1.V.SDT_w$CR[i]))
        }
}

# calculate the d prime for each condition
df.M1.V.SDT_w$dprime <- mapply(dprime,df.M1.V.SDT_w$hitR,df.M1.V.SDT_w$faR)

# transfor from long format to wide format
df.M1.V.SDT_ww <- dcast(df.M1.V.SDT_w, Subject ~ Identity + Morality ,value.var = "dprime")

# anova for d prime with 2*2 design
e7_L.d_anova <- ezANOVA(df.M1.V.SDT_w,dv = dprime, wid = Subject, within=.(Morality,Identity), type=3)

df.M1.V.SDT_w$Subject <- factor(df.M1.V.SDT_w$Subject)  # make the subject "factor"

# using Bayesfactor analysis (which is not reported in the document)
e7_L.dprime_anova_bf  <- anovaBF(dprime ~ Morality*Identity + Subject, data = df.M1.V.SDT_w, whichRandom = "Subject")

# Good self vs Bad self
e7_L.d.t.good_bad_slf <- t.test(df.M1.V.SDT_ww$Self_Good,df.M1.V.SDT_ww$Self_Bad,paired = TRUE)     # t-test
df.M1.V.SDT_ww$good_bad_slf <- df.M1.V.SDT_ww$Self_Good - df.M1.V.SDT_ww$Self_Bad                   # calculate the differences
e7_L.d.t.good_bad_slf.CI <- bootES(df.M1.V.SDT_ww$good_bad_slf,R = 20000, effect.type = "cohens.d") # calculate the cI using BootES

# give new names for each key information, for later use in the markdown
e7_L.d.tvalue.good_bad_slf <- round(as.numeric(e7_L.d.t.good_bad_slf[[1]]),3)
e7_L.d.df.M1.good_bad_slf <- as.numeric(e7_L.d.t.good_bad_slf[[2]])
e7_L.d.pvalue.good_bad_slf.adj <- p.adjust(as.numeric(e7_L.d.t.good_bad_slf[[3]]),"bonferroni",4)
e7_L.d.cohens.good_bad_slf <- round(e7_L.d.t.good_bad_slf.CI[[1]],4) 
e7_L.d.CI.L.good_bad_slf <- round(e7_L.d.t.good_bad_slf.CI[[12]][1],4)
e7_L.d.CI.H.good_bad_slf <- round(e7_L.d.t.good_bad_slf.CI[[12]][2],4)

# Good other vs. Bad other
e7_L.d.t.good_bad_oth <- t.test(df.M1.V.SDT_ww$Other_Good,df.M1.V.SDT_ww$Other_Bad,paired = TRUE)
df.M1.V.SDT_ww$good_bad_oth <- df.M1.V.SDT_ww$Other_Good - df.M1.V.SDT_ww$Other_Bad
e7_L.d.t.good_bad_oth.CI <- bootES(df.M1.V.SDT_ww$good_bad_oth,R = 20000, effect.type = "cohens.d")

e7_L.d.tvalue.good_bad_oth  <- round(as.numeric(e7_L.d.t.good_bad_oth[[1]]),3)
e7_L.d.df.M1.good_bad_oth  <- as.numeric(e7_L.d.t.good_bad_oth[[2]])
e7_L.d.pvalue.good_bad_oth.adj <- p.adjust(as.numeric(e7_L.d.t.good_bad_oth[[3]]),"bonferroni",4)
e7_L.d.cohens.good_bad_oth <- round(e7_L.d.t.good_bad_oth.CI[[1]],4) 
e7_L.d.CI.L.good_bad_oth <- round(e7_L.d.t.good_bad_oth.CI[[12]][1],4)
e7_L.d.CI.H.good_bad_oth <- round(e7_L.d.t.good_bad_oth.CI[[12]][2],4)

# Good self vs Good other
e7_L.d.t.slf_oth_good <- t.test(df.M1.V.SDT_ww$Self_Good,df.M1.V.SDT_ww$Other_Good,paired = TRUE)
df.M1.V.SDT_ww$slf_oth_good <- df.M1.V.SDT_ww$Self_Good - df.M1.V.SDT_ww$Other_Good
e7_L.d.t.slf_oth_good.CI <- bootES(df.M1.V.SDT_ww$slf_oth_good,R = 20000, effect.type = "cohens.d")

e7_L.d.tvalue.slf_oth_good  <- round(as.numeric(e7_L.d.t.slf_oth_good[[1]]),3)
e7_L.d.df.M1.slf_oth_good  <- as.numeric(e7_L.d.t.slf_oth_good[[2]])
e7_L.d.pvalue.slf_oth_good.adj <- p.adjust(as.numeric(e7_L.d.t.slf_oth_good[[3]]),"bonferroni",4)
e7_L.d.cohens.slf_oth_good <- round(e7_L.d.t.slf_oth_good.CI[[1]],4) 
e7_L.d.CI.L.slf_oth_good <- round(e7_L.d.t.slf_oth_good.CI[[12]][1],4)
e7_L.d.CI.H.slf_oth_good <- round(e7_L.d.t.slf_oth_good.CI[[12]][2],4)

# Bad self vs. Bad other
e7_L.d.t.slf_oth_bad <- t.test(df.M1.V.SDT_ww$Self_Bad,df.M1.V.SDT_ww$Other_Bad,paired = TRUE)
df.M1.V.SDT_ww$slf_oth_bad <- df.M1.V.SDT_ww$Self_Bad - df.M1.V.SDT_ww$Other_Bad
e7_L.d.t.slf_oth_bad.CI <- bootES(df.M1.V.SDT_ww$slf_oth_bad,R = 20000, effect.type = "cohens.d")
